---
title: "Lab: Logstic Regression"
author: "Hicham Zmarrou"
date: "`r Sys.Date()`"
output:
  html_notebook:
    highlight: pygments
    number_sections: no
    theme: cosmo
    toc: yes
    toc_float: yes
  html_document:
    df_print: paged
    toc: yes
  word_document:
    toc: yes
---
___________________________________________________________________________________________________________


```{r global_options, include=FALSE}
knitr::opts_chunk$set(fig.width=12, fig.height=8, fig.path='img/', echo=TRUE, warning=FALSE, message=FALSE,
                      digits = 3)
```


We will begin by examining some numerical and graphical summaries of
the Smarket data, which is part of the `ISLR` library. This data set consists of
percentage returns for the `S&P 500` stock index over 1, 250 days, from the
beginning of 2001 until the end of 2005. For each date, we have recorded
the percentage returns for each of the five previous trading days, `Lag1`
through `Lag5`. We have also recorded Volume (the number of shares traded) on the previous day, in billions), Today (the percentage return on the date
in question) and Direction (whether the market was Up or Down on this
date).

```{r}
library (ISLR)
names(Smarket)
dim(Smarket)
summary(Smarket)
help(Smarket)
```


Next, we will fit a logistic regression model in order to predict `Direction`
using `Lag1` through `Lag5` and `Volume`. The `glm()` function fits generalized
`glm()` linear models, a class of models that includes logistic regression. The syntax
generalized of the `glm()` function is similar to that of `lm()`, except that we must pass in linear model the argument `family=binomial` in order to tell `R` to run a logistic regression
rather than some other type of generalized linear model.

```{r}
glm.fits <- glm(Direction ~ Lag1+Lag2+Lag3+Lag4+Lag5+Volume,data=Smarket ,family =binomial )
summary (glm.fits)
```
The smallest `p-value` here is associated with `Lag1`. The negative coefficient
for this predictor suggests that if the market had a positive return yesterday,
then it is less likely to go up today. However, at a value of 0.15, the p-value
is still relatively large, and so there is no clear evidence of a real association
between Lag1 and Direction.


We use the coef() function in order to access just the coefficients for this
fitted model. We can also use the summary() function to access particular
aspects of the fitted model, such as the p-values for the coefficients.

```{r}
coef(glm.fits)
summary(glm.fits)$coef
```

The `predict()` function can be used to predict the probability that the
market will go up, given values of the predictors. The `type="response"`
option tells R to output probabilities of the form `P(Y = 1|X)`, as opposed
to other information such as the logit. If no data set is supplied to the
`predict()` function, then the probabilities are computed for the training
data that was used to fit the logistic regression model. Here we have printed
only the first ten probabilities. We know that these values correspond to
the probability of the market going up, rather than down, because the
contrasts() function indicates that R has created a dummy variable with
a 1 for Up.

```{r}
attach(Smarket)
contrasts(Direction)

glm.probs <- predict(glm.fits, type = "response")
glm.probs [1:10]
```

In order to make a prediction as to whether the market will go up or down on a particular day, we must convert these predicted probabilities into class labels, `Up` or `Down`. The following two commands create a vector of class predictions based on whether the predicted probability of a market increase is greater than or less than 0.5.


```{r}
glm.pred=rep("Down " ,1250)
glm.pred[glm.probs >.5]=" Up"
```

The first command creates a vector of 1,250 Down elements. The second line
transforms to Up all of the elements for which the predicted probability of a
market increase exceeds 0.5. Given these predictions, the `table()` function
can be used to produce a confusion matrix in order to determine how many
observations were correctly or incorrectly classified.

```{r}
table(glm.pred ,Direction )
mean(glm.pred== Direction )
```

```{r}
train <- (Year <2005)
Smarket.2005 <- Smarket[!train,]
dim(Smarket.2005)
Direction.2005 <-  Direction[!train]
```

We now fit a logistic regression model using only the subset of the observations
that correspond to dates before 2005, using the subset argument.
We then obtain predicted probabilities of the stock market going up for
each of the days in our test set-that is, for the days in 2005.


```{r}

glm.fits  <- glm(Direction ~ Lag1+Lag2+Lag3+Lag4+Lag5+Volume , data=Smarket ,family =binomial ,subset = train)
glm.probs <- predict(glm.fits, Smarket.2005 , type = "response")

```

Notice that we have trained and tested our model on two completely separate
data sets: training was performed using only the dates before 2005,
and testing was performed using only the dates in 2005. Finally, we compute
the predictions for 2005 and compare them to the actual movements
of the market over that time period.

```{r}
glm.pred=rep ("Down" ,252)
glm.pred[glm.probs >.5]="Up"
table(glm.pred,Direction.2005)
mean(glm.pred == Direction.2005)
mean(glm.pred != Direction.2005)
```

Below we have refit the logistic regression using just `Lag1` and `Lag2`, which seemed to have the highest predictive power in
the original logistic regression model.

```{r}
glm.fits=glm(Direction~Lag1+Lag2 ,data=Smarket ,family =binomial , subset =train)

glm.probs =predict (glm.fits,Smarket.2005 , type="response")
glm.pred=rep ("Down" ,252)
glm.pred[glm.probs >.5]= "Up"
table(glm.pred ,Direction.2005)
mean(glm.pred== Direction.2005)
```

Suppose that we want to predict the returns associated with particular values of `Lag1` and `Lag2`. In particular, we want to predict Direction on a day when `Lag1` and `Lag2` equal 1.2 and 1.1, respectively, and on a day when they equal 1.5 and ~0.8. We do this using the `predict()` function.


```{r}
predict (glm.fits,newdata =data.frame(Lag1=c(1.2 ,1.5) , Lag2=c(1.1 , -0.8) ),type = "response")
```
