---
title: "Lab: Linear Regression"
author: "Hicham Zmarrou"
date: "`r Sys.Date()`"
output:
  html_notebook:
    highlight: pygments
    number_sections: no
    theme: cosmo
    toc: yes
    toc_float: yes
  html_document:
    df_print: paged
    toc: yes
  word_document:
    toc: yes
---
___________________________________________________________________________________________________________


The MASS library contains the Boston data set, which records medv (median
house value) for 506 neighborhoods around Boston. We will seek to predict
medv using 13 predictors such as rm (average number of rooms per house),
age (average age of houses), and lstat (percent of households with low
socioeconomic status).

```{r}
library(MASS)
data(Boston)
names(Boston)
```

To find out more about the data set, we can type ?Boston.
We will start by using the lm() function to fit a simple linear regression lm() model, with medv as the response and lstat as the predictor. The basic
syntax is lm(y∼x,data), where y is the response, x is the predictor, and
data is the data set in which these two variables are kept.


```{r}
lm.fit=lm(medv∼lstat)
```
The command causes an error because R does not know where to find
the variables medv and lstat. The next line tells R that the variables are
in Boston. If we attach Boston, the first line works fine because R now
recognizes the variables.

```{r}
lm.fit=lm(medv∼lstat , data=Boston)
attach(Boston)
lm.fit=lm(medv∼lstat)
```

If we type `lm.fit`, some basic information about the model is output.
For more detailed information, we use `summary(lm.fit)`. This gives us `p-values`
and standard errors for the coefficients, as well as the `R^2` statistic
and `F-statistic`  for the model.

